---
title: <span style="color:red">SQL</span>
output: html_document
---
```{r warning=FALSE,message=FALSE}
library(Lahman)
library(sqldf)
```

###<span style="color:red">Filtering Rows and Columns</span>

>In SQL you can filter columns by using SELECT and rows by using WHERE.  Let's look at an example using the Lahman baseball database.  First, we should load the Lahman package.  We will also need the sqldf package.

```{r message=FALSE, warning=FALSE, comment=NA}
library(Lahman)
library(sqldf)
```

>Let's take a look at an example.  Suppose we want to see the home run totals for the 1927 Yankees.  We could write:

```{r message=FALSE, warning=FALSE, comment=NA}
query<-"SELECT playerID,teamID,yearID,HR FROM Batting WHERE teamID='NYA' and yearID=1927"
sqldf(query)
```
#------------------------------------------
###<span style="color:red">NY Yankees with 40 or more Homeruns</span>
>Moving on we can query out which NY Yankees players hit 40 or more homeruns with the following code:

```{r message=FALSE, warning=FALSE, comment=NA}
query<-"SELECT playerID,yearID,teamID,HR FROM Batting WHERE teamID='NYA' and HR>=40"
sqldf(query)
```
#------------------------------------------
###<span style="color:red">Strike Outs and Homeruns</span>
>Adding the Strike Outs into the query we can further filter out which players hit over 40 homeruns AND had less than 60 strike outs with the following:

```{r message=FALSE, warning=FALSE, comment=NA}
query<-"SELECT playerID,teamID,yearID,HR,SO FROM Batting WHERE HR>40 and SO<60"
sqldf(query)
```
#------------------------------------------
###<span style="color:red">Team Filter</span>
>Finally we're filtering all the teams to just the Phillies, query the years in the 1970's (using a range), and finally extract the players with more than 30 homeruns.

```{r message=FALSE, warning=FALSE, comment=NA}
query<-"SELECT playerID,teamID,yearID,HR FROM Batting WHERE teamID='PHI' and yearID>=1970 and yearID<=1979 and HR>30"
sqldf(query)
```
#------------------------------------------
###<span style="color:red">ORDER BY</span>

>This filter extracts every instance of a player hitting more than 50 homeruns.  Furthermore, let's have the players with the highest homerun total at the top.

```{r message=FALSE, warning=FALSE, comment=NA}
query<-"SELECT playerID,yearID,teamID,HR FROM Batting 
WHERE HR>50 
ORDER BY HR DESC"
sqldf(query)
```
#------------------------------------------
###<span style="color:red">Filtering Strike Outs (Less than 10) and At Bat (400 or more)</span>

>This filter extracts every instance of a player strike out less than 10 times; however, they need to be at bat atleast 400 times.

```{r message=FALSE, warning=FALSE, comment=NA}
query<-"SELECT playerID,yearID,teamID,SO,HR,AB FROM Batting 
WHERE SO<10 and AB>=400
ORDER BY SO DESC"
sqldf(query)
```
#------------------------------------------
###<span style="color:red">Group By Filter</span>

>We will be grouping (summating) all of Babe Ruth's career homeruns using the group by function.

```{r message=FALSE, warning=FALSE, comment=NA}
query<-"SELECT playerID, sum(HR) FROM Batting 
WHERE playerID='ruthba01'
GROUP BY playerID"
sqldf(query)
```
#------------------------------------------
###<span style="color:red">Summing all players with Homeruns over 600 in descending order</span>

>We will be grouping (summating) all of a players career homeruns using the group by function and place it in descending order.

```{r message=FALSE, warning=FALSE, comment=NA}
query<-"SELECT playerID, sum(HR) FROM Batting 
GROUP BY playerID
HAVING sum(HR)>=600
ORDER BY HR DESC"
sqldf(query)
```
#------------------------------------------
###<span style="color:red">Averaging Homeruns</span>

>We will be grouping all the players with average homeruns greater than 30 in descending order.

```{r message=FALSE, warning=FALSE, comment=NA}
query<-"SELECT playerID, avg(HR) FROM Batting 
GROUP BY playerID
HAVING avg(HR)>30
ORDER BY avg(HR) DESC"
sqldf(query)
```

#------------------------------------------
###<span style="color:red">First and Last Name</span>

>First and last name listing for Babe Ruth with concatennated tables.

```{r message=FALSE, warning=FALSE, comment=NA}
query<-"SELECT nameFirst,nameLast,teamID,yearID,HR FROM Batting
INNER JOIN Master
ON Batting.playerID=Master.playerID
WHERE Batting.playerID='ruthba01'"
sqldf(query)
```

#------------------------------------------
###<span style="color:red">Over 50 Homeruns with Names and Years</span>

>All instances of players hitting more than 50 homeruns, listed by first and last name, team, year, and homeruns.

```{r message=FALSE, warning=FALSE, comment=NA}
query<-"SELECT nameFirst,nameLast,teamID,yearID,HR FROM Batting
INNER JOIN Master
ON Batting.playerID=Master.playerID
WHERE HR>50
ORDER BY HR DESC"
sqldf(query)
```
#------------------------------------------
###<span style="color:red">Babe Ruth by Team, Year, and Homeruns</span>

>All instances of Babe Ruth by team, year, and homeruns, concatennating two different tables.

```{r message=FALSE, warning=FALSE, comment=NA}
query<-"SELECT playerID,name,Batting.yearID,Batting.HR
FROM Batting INNER JOIN Teams
ON Batting.teamID=Teams.teamID and Batting.yearID=Teams.yearID
WHERE playerID='ruthba01'"
sqldf(query)
```

#------------------------------------------
###<span style="color:red">Babe Ruth by Name and Teams Played in</span>

In this query we concatennate multiple tables and are able to get Babe Ruth's name instead of a code representing Babe Ruth.  We are also able to do the same with all of the teams he has been on.
```{r message=FALSE, warning=FALSE, comment=NA}
query<-"SELECT nameFirst,nameLast,name,Batting.yearID,Batting.HR
FROM (Batting INNER JOIN Master
ON Batting.playerID=Master.playerID) INNER JOIN Teams
ON Batting.teamID=Teams.teamID AND Batting.yearID=Teams.yearID
WHERE Batting.playerID='ruthba01'"

sqldf(query)
```

#------------------------------------------------------------------

###<span style="color:red">All Players with over 600 Career Homeruns in Descending Order</span>
>For this query we are grouping all players with over 600 career homeruns, and put them in descending order.

```{r message=FALSE, warning=FALSE, comment=NA}
query<-"SELECT nameFirst,nameLast,Sum(HR)
FROM Batting INNER JOIN Master
ON Batting.playerID=Master.playerID
GROUP BY Batting.playerID
HAVING Sum(Batting.HR)>=600
ORDER BY Sum(Batting.HR) DESC"

sqldf(query)
```
#------------------------------------------------------------------

###<span style="color:red">All Players with an Average Salary over one million USD...and named Bob</span>
>This is an interesting query where we concatennate two tables using the Salaries table for the first time.  We get all players averaging over one million USD.  The real kicker is that we are also filtering out all those without the name "Bob".

```{r message=FALSE, warning=FALSE, comment=NA}
query<-"SELECT nameFirst,nameLast,avg(salary)
FROM Salaries INNER JOIN Master
ON Salaries.playerID=Master.playerID
WHERE nameFirst='Bob'
GROUP BY Salaries.playerID
HAVING Avg(Salary)>1000000
ORDER BY Avg(Salary) DESC"
sqldf(query)
```